Refinements for open automata

### 2022-09-01 2021-Refinement/paper/refinementarticle.tex
FACS proofs

We thus define the following property that can be understood as follows. Take a specification automaton and its implementation. Consider two related states for the refinement relation. Suppose an identical transition exist in both automata (with same guards and same label), then either the reached states $s_1'$ and $s_2'$ are related by refinement, or there exists another transition of the specification that is also identical and leads to a state $s_1''$ that is related to $s_2'$ in the refinement relation.
Finally, we should take into account that 1) instead of equality, we should check if there is a state such that both transitions are feasible (the conjunction of guards is satisfiable), and that  2) instead of finding another transition of the specification we might find a family of transitions that cover all the cases of the transition of the implementation. This leads us to the following definition.
\begin{definition}[Preservation along identical transitions]\label{def:preservation} \\
Let \(\OAg[1]\) and \(\OAg[2]\) be two OAs.
A relation \(R: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\) between  \(\OAg[1]\) and \(\OAg[2]\)
  is preserved along identical transitions if  it satisfies the following:
% TODO: isn't it a2 = a1x instead of a1x = a1
% TODO: ANS:  we have a1 = a2 so whatever, it is the same, we can change if it is more ``natural''
\begin{multline*}
\forall \mpar{s_1, s_2} \in S_1 \times S_2, \bigsymb{\forall} \mpar{t_1, t_2} = \mpar{\OTx{1}{}{1}{1}, \OTx{2}{}{2}{2}} \in \fOT{s_1} \times \fOT{s_2} \\ 
\bigsymb{\bigexists} 
\mpar{t_{1x} \!=\! \OTx{1}{x}{1x}{1x} \!\in\! \fOT{s_1}}^{x \in X}\!,
 V_1 \uplus V_2 \uplus \biguplus_{x\in X}\!\fvars{t_{1x}} \uplus \fvars{t_2} \uplus \fvars{t_1} \vdash \hspace{6cm}\\  R\mpar{s_1, s_2} \wedge 	g_1 \wedge g_2 \wedge \alpha_1 = \alpha_2 
%\wedge \bigwedge_\subbox{j \in J'_1 \cap J'_2} \beta_{1j} = \beta_{2j} 
\implies \\
%\exists
%\OTx{1}{a}{3}{3}\in \fOT{s_1}. \\
\bigsymb{\bigwedge_{x\in X}}
\mpar{
\alpha_1=\alpha_{1 x}  \land J'_{1x}= J'_2 \land \bigwedge_\subbox{j \in J'_2 } \beta_{2j} = \beta_{1xj}\land g_{1 x}\land
 R\mpar{s'_{1 x}, s'_2}\psubst{\psi_{1 x} \uplus \psi_2}}\\
\end{multline*}
\end{definition}

@@

\item  Finally, we have to prove the following holds:
\begin{multline*}
\bigforall \mpar{s_1, s_3} \in S_1 \times S_3, \bigsymb{\forall} \mpar{t_1, t_3} = \mpar{\OTx{1}{}{1}{1}, \OTx{3}{}{3}{3}} \in \fOT{s_1} \times \fOT{s_3} \\ 
\bigsymb{\bigexists} 
\mpar{t_{1x} = \OTx{1}{x}{1x}{1x} \in \fOT{s_1}}^{x \in X},
 V_1 \uplus V_3\uplus \biguplus_{x\in X}\fvars{t_{1x}} \uplus \fvars{t_1} \uplus \fvars{t_3} \vdash \hspace{6cm}\\  R_{13}\mpar{s_1, s_3} \wedge 	g_1 \wedge g_3 \wedge \alpha_1 = \alpha_3 
%\wedge \bigwedge_\subbox{j \in J'_1 \cap J'_3} \beta_{1j} = \beta_{3j} 
\implies \hspace{6cm}\\
%\exists
%\OTx{1}{a}{3}{3}\in \fOT{s_1}. \\
\bigsymb{\bigwedge_{x\in X}}
\mpar{
\alpha_1=\alpha_{1 x}  \land J'_{1x}= J'_3 \land \bigwedge_\subbox{j \in J'_3 } \beta_{3j} = \beta_{1xj}\land g_{1 x}\land
 R_{13}\mpar{s'_{1 x}, s'_3}\psubst{\psi_{1 x} \uplus \psi_3}}
\end{multline*}

@@

\newpage
\TODO{[RABEA] proof}
Let's consider any $s_{2p} \in (s_{2p})^{p\in P}$. According to the refinement relation between  $A_1$ and $A_2$, we have  the following formula that we denote ($*$):\\
for any open transition $t_1$ in $T_1$  originating from $s_1$:
\begin{align*}
\OTx{1}{}{1}{1} \in \fOT{s_1}
\end{align*}
there exists an indexed family of OTs originating from $s_{2p}$: 
\begin{align*}
\mpar{\OTx{2p}{x}{2px}{2px} \in \fOT{s_{2p}}}^{x \in X_p} 
\end{align*}
such that $\forall x \in X, J'_{2px} \cap H = J'_1 \cap H$ and
\begin{multline*}
%\everymath{\displaystyle}
V_1 \uplus V_2 \uplus \fvars{t_1} \vdash\\ \mpar{R_{12}\mpar{s_1, s_{2p}} \wedge g_1 \implies \operatorname*{\bigsymb{\bigvee}}_{x \in X_p} \mpar{\everymath{\displaystyle}
\begin{array}{l}
			\alpha_1 = \alpha_{2px} \wedge \bigwedge_\subbox{j \in J'_{2px} \cap H} \beta_{1j} = \beta_{2pxj} \nwedge\\
			 g_{2px} \wedge R_{12}\mpar{s'_1, s'_{2px}}\psubst{\psi_1 \uplus \psi_{2px}}
		\end{array}}} 
\end{multline*}	
On the other hand, according to the refinement relation between   $A_2$ and $A_3$, we know condition 3 is satisfied, $\bigforall \mpar{s_{2p}, s_3} \in S_1 \times S_3$ and $\bigsymb{\forall} \mpar{t_{2p}, t_3} \in \fOT{s_{2p}} \times \fOT{s_3}$ and in particular for  the set $(t_{2px})^{x\in X_p}$ so we  have the following that we denote ($**$):
\begin{multline*}
\bigsymb{\forall} \mpar{t_{2px}, t_3} = \mpar{\OTx{2px}{}{2px}{2px}, \OTx{3}{}{3}{3}} \in \fOT{s_{2p}} \times \fOT{s_3} \\ 
\bigsymb{\bigexists} 
\mpar{t_{2pxy} = \OTx{2px}{y}{2pxy}{2pxy} \in \fOT{s_{2p}}}^{y \in Y_X},
V_2 \uplus V_3\uplus \biguplus_{y\in Y}\fvars{t_{2pxy}} \uplus \fvars{t_3} \uplus \fvars{t_{2p}} \vdash \hspace{6cm}\\  R_{23}\mpar{s_{2p}, s_3} \wedge 	g_{2px}\wedge g_3 \wedge \alpha_{2px}= \alpha_3 
%\wedge \bigwedge_\subbox{j \in J'_2 \cap J'_3} \beta_{2pj} = \beta_{3j} 
\implies \hspace{6cm}\\
%\exists
%\OTx{1}{a}{3}{3}\in \fOT{s_2}. \\
\bigsymb{\bigwedge_{y\in Y_X}}
\mpar{
\alpha_{2px}=\alpha_{2p xy}  \land J'_{2pxy}= J'_3 \land \bigwedge_\subbox{j \in J'_3 } \beta_{3j} = \beta_{2pxyj}\land g_{2p xy}\land
 R_{23}\mpar{s'_{2p xy}, s'_3}\psubst{\psi_{2p xy} \uplus \psi_3}}\\
\end{multline*}
Again according to the refinement  relation between  $A_1$ and $A_2$ and condition 3  the following is satisfied $\bigforall \mpar{s_1, s_{2p}} \in S_1 \times S_2$ , in particular over the family of transitions in $(t_{2pxy})^{y\in Y_X}$:
\begin{multline*}
\bigsymb{\forall} \mpar{t_1, t_{2pxy}} = \mpar{\OTx{1}{}{1}{1}, \OTx{2pxy}{}{2pxy}{2pxy}} \in \fOT{s_1} \times \fOT{s_{2p}} \\ 
\bigsymb{\bigexists} 
\mpar{t_{1z} = \OTx{1}{z}{1z}{1z} \in \fOT{s_1}}^{z \in Z},
 V_1 \uplus V_2 \uplus \biguplus_{x\in X}\fvars{t_{1x}} \uplus \fvars{t_1} \uplus \fvars{t_{2p}} \vdash \hspace{6cm}\\  R_{12}\mpar{s_1, s_{2p}} \wedge 	g_1 \wedge g_{2pxy} \wedge \alpha_1 = \alpha_{2pxy} 
%\wedge \bigwedge_\subbox{j \in J'_1 \cap J'_{2p}} \beta_{1j} = \beta_{2pxyj} 
\implies \hspace{6cm}\\
%\exists
%\OTx{1}{a}{3}{3}\in \fOT{s_1}. \\
\bigsymb{\bigwedge_{z\in Z}}
\mpar{
\alpha_1=\alpha_{1 z}  \land J'_{1z}= J'_{2pxy} \land \bigwedge_\subbox{j \in J'_{2pxy} } \beta_{2pxyj} = \beta_{1zj}\land g_{1 z}\land
 R_{12}\mpar{s'_{1 z}, s'_{2pxy}}\psubst{\psi_{1 z} \uplus \psi_{2pxy}}}
\end{multline*}

>>>>

We thus define the following property that can be understood as follows. Take a specification automaton and its implementation. Consider two related states for the refinement relation. Suppose an identical transition exist in both automata (with same guards and same label), then either the reached states $s_1'$ and $s_2'$ are related by refinement, or there exists another transition of the specification that is also identical and leads to a state $s_1''$ that is related to $s_2'$ in the refinement relation.
Finally, we should take into account that 1) instead of equality, we should check if there is a state such that both transitions are feasible (the conjunction of guards is satisfiable), and that  2) instead of finding another transition of the specification we might find a family of transitions that cover all the cases of the transition of the implementation. This leads us to the following definition.
\begin{definition}[Preservation along identical transitions]\label{def:preservation} \\
Let \(\OAg[1]\) and \(\OAg[2]\) be two OAs.
A relation \(R: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\) between  \(\OAg[1]\) and \(\OAg[2]\)
  is preserved along identical transitions if  it satisfies the following:
% TODO: isn't it a2 = a1x instead of a1x = a1
% TODO: ANS:  we have a1 = a2 so whatever, it is the same, we can change if it is more ``natural''
\begin{multline*}
\forall \mpar{s_1, s_2} \in S_1 \times S_2, \bigsymb{\forall} \mpar{t_1, t_2} = \mpar{\OTx{1}{}{1}{1}, \OTx{2}{}{2}{2}} \in \fOT{s_1} \times \fOT{s_2} \\ 
\bigsymb{\bigexists} 
\mpar{t_{1x} \!=\! \OTx{1}{x}{1x}{1x} \!\in\! \fOT{s_1}}^{x \in X}\!,
 V_1 \uplus V_2 \uplus \biguplus_{x\in X}\!\fvars{t_{1x}} \uplus \fvars{t_2} \uplus \fvars{t_1} \vdash \hspace{6cm}\\  R\mpar{s_1, s_2} \wedge 	g_1 \wedge g_2 \wedge \alpha_1 = \alpha_2 
%\wedge \bigwedge_\subbox{j \in J'_1 \cap J'_2} \beta_{1j} = \beta_{2j} 
\implies \\
%\exists
%\OTx{1}{a}{3}{3}\in \fOT{s_1}. \\
\bigsymb{\bigvee_{x\in X}}
\mpar{
\alpha_1=\alpha_{1 x}  \land J'_{1x}= J'_2 \land \bigwedge_\subbox{j \in J'_2 } \beta_{2j} = \beta_{1xj}\land g_{1 x}\land
 R\mpar{s'_{1 x}, s'_2}\psubst{\psi_{1 x} \uplus \psi_2}}\\
\end{multline*}
\end{definition}

@@

\item  Finally, we have to prove the following holds:
\begin{multline*}
\bigforall \mpar{s_1, s_3} \in S_1 \times S_3, \bigsymb{\forall} \mpar{t_1, t_3} = \mpar{\OTx{1}{}{1}{1}, \OTx{3}{}{3}{3}} \in \fOT{s_1} \times \fOT{s_3} \\ 
\bigsymb{\bigexists} 
\mpar{t_{1x} = \OTx{1}{x}{1x}{1x} \in \fOT{s_1}}^{x \in X},
 V_1 \uplus V_3\uplus \biguplus_{x\in X}\fvars{t_{1x}} \uplus \fvars{t_1} \uplus \fvars{t_3} \vdash \hspace{6cm}\\  R_{13}\mpar{s_1, s_3} \wedge 	g_1 \wedge g_3 \wedge \alpha_1 = \alpha_3 
%\wedge \bigwedge_\subbox{j \in J'_1 \cap J'_3} \beta_{1j} = \beta_{3j} 
\implies \hspace{6cm}\\
%\exists
%\OTx{1}{a}{3}{3}\in \fOT{s_1}. \\
\bigsymb{\bigvee_{x\in X}}
\mpar{
\alpha_1=\alpha_{1 x}  \land J'_{1x}= J'_3 \land \bigwedge_\subbox{j \in J'_3 } \beta_{3j} = \beta_{1xj}\land g_{1 x}\land
 R_{13}\mpar{s'_{1 x}, s'_3}\psubst{\psi_{1 x} \uplus \psi_3}}
\end{multline*}

@@

Consider $(t_1,t_3)\in\fOT{s_1} \times \fOT{s_3}$

\newpage
\TODO{[RABEA] proof}
Let's consider any $s_{2p} \in (s_{2p})^{p\in P}$. Consider  $t_1$ in $T_1$  originating from $s_1$:
\begin{align*}
\OTx{1}{}{1}{1} \in \fOT{s_1}
\end{align*}
According to the refinement relation between  $A_1$ and $A_2$ there exists an indexed family of OTs originating from $s_{2p}$: 
\begin{align*}
\mpar{\OTx{2p}{x}{2px}{2px} \in \fOT{s_{2p}}}^{x \in X_p} 
\end{align*}
such that $\forall x \in X, J'_{2px} \cap H = J'_1 \cap H$ and
\begin{multline}\label{lab1}\tag{$*$}
%\everymath{\displaystyle}
V_1 \uplus V_2 \uplus \fvars{t_1} \vdash\\ \mpar{R_{12}\mpar{s_1, s_{2p}} \wedge g_1 \implies \operatorname*{\bigsymb{\bigvee}}_{x \in X_p} \mpar{\everymath{\displaystyle}
\begin{array}{l}
			\alpha_1 = \alpha_{2px} \wedge \bigwedge_\subbox{j \in J'_{2px} \cap H} \beta_{1j} = \beta_{2pxj} \nwedge\\
			 g_{2px} \wedge R_{12}\mpar{s'_1, s'_{2px}}\psubst{\psi_1 \uplus \psi_{2px}}
		\end{array}}} 
\end{multline}	
On the other hand, according to the refinement relation between   $A_2$ and $A_3$, we know condition 3 is satisfied, $\bigforall \mpar{s_{2p}, s_3} \in S_2 \times S_3$ and $\bigsymb{\forall} \mpar{t_{2p}, t_3} \in \fOT{s_{2p}} \times \fOT{s_3}$ and in particular for  the set $(t_{2px})^{x\in X_p}$ so we  have the following that we denote ($**$):
\begin{multline*}
\bigsymb{\forall} \mpar{t_{2px}, t_3} = \mpar{\OTx{2px}{}{2px}{2px}, \OTx{3}{}{3}{3}} \in \fOT{s_{2p}} \times \fOT{s_3} \\ 
\bigsymb{\bigexists} 
\mpar{t_{2pxy} = \OTx{2px}{y}{2pxy}{2pxy} \in \fOT{s_{2p}}}^{y \in Y_X},
V_2 \uplus V_3\uplus \biguplus_{y\in Y}\fvars{t_{2pxy}} \uplus \fvars{t_3} \uplus \fvars{t_{2p}} \vdash \hspace{6cm}\\  R_{23}\mpar{s_{2p}, s_3} \wedge 	g_{2px}\wedge g_3 \wedge \alpha_{2px}= \alpha_3 
%\wedge \bigwedge_\subbox{j \in J'_2 \cap J'_3} \beta_{2pj} = \beta_{3j} 
\implies \hspace{6cm}\\
%\exists
%\OTx{1}{a}{3}{3}\in \fOT{s_2}. \\
\bigsymb{\bigvee_{y\in Y_X}}
\mpar{
\alpha_{2px}=\alpha_{2p xy}  \land J'_{2pxy}= J'_3 \land \bigwedge_\subbox{j \in J'_3 } \beta_{3j} = \beta_{2pxyj}\land g_{2p xy}\land
 R_{23}\mpar{s'_{2p xy}, s'_3}\psubst{\psi_{2p xy} \uplus \psi_3}}\\
\end{multline*}
combining $(*)$ and $(**)$ with the definition of the set $s_{2p}$, we obtain (variables on the left  of the $\vdash$ omitted for readability)
\begin{align*}
 R_{13}\mpar{s_1, s_3} &\wedge 	g_1 \wedge g_3 \wedge \alpha_1 = \alpha_3 \implies \\
&(\bigvee_{{p\in P}}\mpar{R_{12}\mpar{s_1,s_{2p}} \wedge R_{23}\mpar{s_{2p},s_3}}) \wedge 	g_1 \wedge g_3 \wedge \alpha_1 = \alpha_3  \implies \\
& (\bigvee_{{p\in P}}{\bigsymb{\bigvee}}_{x \in X_p} \left(\mpar{\everymath{\displaystyle}
\begin{array}{l}
			\alpha_1 = \alpha_{2px}= \alpha_3 \wedge \bigwedge_\subbox{j \in J'_{2px} \cap H} \beta_{1j} = \beta_{2pxj} \nwedge\\
			 g_{2px} \wedge R_{12}\mpar{s'_1, s'_{2px}}\psubst{\psi_1 \uplus \psi_{2px}}
		\end{array}} \right) 
	\wedge R_{23}\mpar{s_{2p},s_3}  \wedge g_3  \implies \\
\end{align*}

Again according to the refinement  relation between  $A_1$ and $A_2$ and condition 3  the following is satisfied $\bigforall \mpar{s_1, s_{2p}} \in S_1 \times S_2$ , in particular over the family of transitions in $(t_{2pxy})^{y\in Y_X}$:
\begin{multline*}
\bigsymb{\forall} \mpar{t_1, t_{2pxy}} = \mpar{\OTx{1}{}{1}{1}, \OTx{2pxy}{}{2pxy}{2pxy}} \in \fOT{s_1} \times \fOT{s_{2p}} \\ 
\bigsymb{\bigexists} 
\mpar{t_{1z} = \OTx{1}{z}{1z}{1z} \in \fOT{s_1}}^{z \in Z},
 V_1 \uplus V_2 \uplus \biguplus_{x\in X}\fvars{t_{1x}} \uplus \fvars{t_1} \uplus \fvars{t_{2p}} \vdash \hspace{6cm}\\  R_{12}\mpar{s_1, s_{2p}} \wedge 	g_1 \wedge g_{2pxy} \wedge \alpha_1 = \alpha_{2pxy} 
%\wedge \bigwedge_\subbox{j \in J'_1 \cap J'_{2p}} \beta_{1j} = \beta_{2pxyj} 
\implies \hspace{6cm}\\
%\exists
%\OTx{1}{a}{3}{3}\in \fOT{s_1}. \\
\bigsymb{\bigwedge_{z\in Z}}
\mpar{
\alpha_1=\alpha_{1 z}  \land J'_{1z}= J'_{2pxy} \land \bigwedge_\subbox{j \in J'_{2pxy} } \beta_{2pxyj} = \beta_{1zj}\land g_{1 z}\land
 R_{12}\mpar{s'_{1 z}, s'_{2pxy}}\psubst{\psi_{1 z} \uplus \psi_{2pxy}}}
\end{multline*}
